DOM Review Summary

### 2024-03-15 Analysis/ReviewSummary/ReviewSummary.qmd
Threatened taxa
Details on threats to taxa and IUCN status of taxa

kable(species_taxaTable)
```

@@

Only approximately half of articles modelled occupancy dynamics for a single species. Others fit multiple occupancy models: either multi-species models where 2+ species were modelled in concert, or many independent models for a suite of species. The largest number of species modelled in a single paper was **`{r} max(species_Number$Taxa_Number)`**.

>>>>

kable(species_taxaTable)
species_taxaTable |>
  mutate(Taxa = fct_reorder(Taxa, Articles)) |>

  ggplot() +
  geom_col(aes(x = Articles, y = Taxa, fill = Taxa)) +
  scale_fill_manual(limits = c("Bird", "Threatened", "Mammal", 
                               "Herptile", "Invasive", "Invertebrate",
                               "Lichen"),
                    values = c("darkslategrey", "firebrick4", "darkslategrey", 
                               "darkslategrey", "cyan4", "darkslategrey", 
                               "darkslategrey")) +
  theme(legend.position = "none") +
  labs(title = "Modelled taxa", y = "", 
       x = "number of articles")

## Threats to modelled taxa

Of papers which describe any threats to their focal taxa, by far the most frequent threat was habitat loss.

```{r Species_Threats}
#| echo: false

species_clean |>
  filter(!Threat_Category %in% c("NA", "None"),
         !is.na(Threat_Category)) |> 
  separate_rows(Threat_Category, sep = ", ") |>
  select(ID_clean, Threat_Category) |>
  distinct() |>
  mutate(Total = length(unique(ID_clean))) |>
  summarise(Articles = n(), 
            Proportion = round(Articles/mean(Total), 2),
            .by = Threat_Category)|>
  mutate(Threat_Category = case_when(Threat_Category == "HabitatLoss" ~ "Habitat loss",
                                     Threat_Category == "ClimateChange" ~ "Climate change",
                                     Threat_Category == "Other_Pollution" ~ "Pollution",
                                     .default = Threat_Category)) |>
  mutate(Threat_Category = fct_reorder(Threat_Category, Articles)) |>

  ggplot() +
  geom_col(aes(x = Proportion, y = Threat_Category), fill = "darkslategrey") +
  xlim(c(0,1)) +
  labs(title = "Threats to study taxa",
       subtitle = "Of articles which explicitly mention any threats",
       x = "Proportion of articles", y = "")
```

```{r Species_IUCN}
#| echo: false

species_clean |>
  summarise(Articles = n(), .by = IUCN_Status) |>
  filter(!IUCN_Status %in% c("Not applicable", "NA")) |>
  mutate(IUCN_Status = case_when(IUCN_Status == "NL" ~ "Not listed",
                                 .default = IUCN_Status)) |>
  mutate(IUCN_Status = fct(IUCN_Status,
                           levels = c("LC", "NT", "VU",
                                      "EN", "CR", "Not listed"))) |>

  ggplot() +
  geom_col(aes(x = IUCN_Status, y = Articles, fill = IUCN_Status)) +
  scale_fill_manual(values = c("darkslategrey", "darkorange3",
                               "darkorange3", "darkorange3",
                               "darkorange3", "darkslategrey")) +
  labs(title = "IUCN Red list status",
       subtitle = "Invasives, higher-order taxa, and grouped species excluded",
       y = "Number of articles", x = "IUCN Red List Status") +
  theme(legend.position = "none")
```

@@

Only approximately half of articles modelled occupancy dynamics for a single taxa. Others fit multiple occupancy models: either multi-species models where 2+ taxa were modelled in concert, or many independent models for a suite of taxa. The largest number of species modelled in a single paper was **`{r} max(species_Number$Taxa_Number)`**.
